### 2020-07-18 article.tex
First attempt at fixing the last (?) issues

% REQUIRED
\begin{abstract}
  Contemporary applications in computational science and engineering often
    require the solution of linear systems which may be of different sizes,
    shapes, and structures. The goal of this paper is to explain how two
    libraries, PETSc and HPDDM, have been interfaced together in order to offer
    end-users robust overlapping Schwarz preconditioners and advanced Krylov
    methods featuring recycling and block solution techniques.
    The flexibility of the implementation %, which for example may deal with various matrix formats,
    is showcased and explained with minimalist, easy-to-run and reproducible examples
    ranging from {\color{red}fracture mechanics to linear stability analysis}
    to ease the integration of these algorithms into more advanced frameworks.
\end{abstract}

@@

\section{Introduction}
Implementing and maintaining robust and scalable applications is highly relevant
in the field of computational science and engineering. The efficient production of
such applications requires a rich ecosystem of state-of-the-art reusable libraries
\cite{Bartlett:2017:XFT}. In this paper, we focus on the interoperability of two such libraries.

@@

% Htool ?
Here, we describe the integration of the HPDDM solvers suite consisting of advanced iterative
methods and robust domain decomposition preconditioners into PETSc.
The design principles of the interface are showcased considering % alongside
minimalist and easy-to-run examples, with a focus on fulfilling reproducibility requirements as advocated by the
Association for Computing Machinery
\url{https://www.acm.org/publications/policies/artifact-review-badging}, as well as to 
ease the integration of these algorithms into more advanced frameworks
and facilitate the analysis and comparison of performance results~\cite{hoefler2015scientific}.

@@

{\color{red} THIS PARAGRAPH IS A BIT DETACHED FROM THE REST OF THE SECTION.
WHAT ABOUT ADDING A SMALL PARAGRAPH IN THE INTRODUCTION (in the spirit of third party code that directly benefits from the PETSc-HPDDM interface) RIGHT BEFORE THE OUTLINE OF THE PAPER
AND CONTINUE THE DISCUSSION IN ~\cref{sec:lobpcg}?}
With these added functions, it is now possible to do code refactorization for
applications that deal with multiple right-hand sides. This is the case of some
eigensolvers in SLEPc~\cite{hernandez2005ssf}, the Scalable Library for
Eigenvalue Problem computations. SLEPc is an add-on library that extends PETSc
with several classes for linear and nonlinear eigenvalue problems, as well as with
tools to facilitate the use of shift-and-invert spectral transformations.
The implementations {\color{red} spell out LOBPGC and CISS} of LOBPCG~\cite{knyazev2001toward} and CISS~\cite{sakurai2003projection}
in SLEPc were previously not using blocking when applying the preconditioner or solving linear systems.
In version 3.14.0, we have adapted these solvers to employ \pk{KSPMatSolve} and \pk{PCMatApply}.
In the case of LOBPCG, which is a purely blocked method, all steps were already
implemented as block operations, except the applications of the preconditioner,
which ruined the benefits of blocking. With the new implementation, the gains
are striking, as reported in~\cref{sec:lobpcg}.

@@

All results have been obtained using double precision arithmetic and 32-bit integers. The scaled efficiency measured in GFLOP per second is reported

@@

\subsubsection{Blocking inside SLEPc\label{sec:lobpcg}}

@@

{\color{red} I WOULD PREFER CONCLUSIONS in ACTIVE TENSE. ALSO, THE TONE IS 
TOO MUCH COLLOQUIAL (i.e. using "and such"). THERE ARE FEW POINTS WE CAN ADD FOR
FUTURE WORK: BDD-GENEO?, GPU AWARE HPDDM? SOME MORE SPECIFICS ON BLOCK KRYLOV-SCHUR?}

In this paper, the interface between PETSc and HPDDM is presented. The new
registered Krylov and overlapping Schwarz methods are explained and numerical
examples are provided to show the applicability of these. Solving well-known
problems, various options have been described. %Some more unconventional test
%cases were commented as well, for example using \pk{MatKAIJ} or \pk{MatIS} matrix types.
Overall, the interface paves the way for having recycling and block Krylov
methods plus robust overlapping Schwarz methods using the GenEO framework in
PETSc.

@@

Concerning HPDDM, its integration inside PETSc offers a much-needed flexibility
compared to the standard implementation. Indeed, it is now possible to switch at
runtime many parameters that are determined at compile-time in the standard implementation, e.g.,
subdomain solvers, matrix types, and such.

Concerning SLEPc, as a future work we plan to implement new solvers (or extend existing
ones) to further exploit the block solve primitives, for instance block Krylov-Schur.

@@

@@

\bibitem{ambartsumyan2020hierarchical}
{\sc I.~Ambartsumyan, W.~Boukaram, T.~Bui-Thanh, O.~Ghattas, D.~Keyes, G.~Stadler, G.~Turkiyyah and S.~Zampini},
  {\em Hierarchical Matrix Approximations of Hessians Arising in Inverse Problems Governed by PDEs},
  arXiv preprint arXiv:2003.10173 (2020).

\bibitem{anderson2019mfem}
{\sc R.~Anderson, J.~Andrej, A.~Barker, J.~Bramwell, J.-S. Camier, J.~Cerveny,
  V.~Dobrev, Y.~Dudouit, A.~Fisher, T.~Kolev, et~al.}, {\em {MFEM: a modular
  finite element methods library}}, arXiv:1911.09220,  (2019),
  \url{http://mfem.org}.

@@

>>>>

% REQUIRED
\begin{abstract}
  Contemporary applications in computational science and engineering often
    require the solution of linear systems which may be of different sizes,
    shapes, and structures. The goal of this paper is to explain how two
    libraries, PETSc and HPDDM, have been interfaced together in order to offer
    end-users robust overlapping Schwarz preconditioners and advanced Krylov
    methods featuring recycling and block solution techniques.
    The flexibility of the implementation %, which for example may deal with various matrix formats,
    is showcased and explained with minimalist, easy-to-run, and reproducible examples,
    to ease the integration of these algorithms into more advanced frameworks.
    The examples simulate the linear or nonlinear solves of typical
    applications in fracture mechanics, linear stability analysis or more
    broadly-speaking eigenanalysis, elasticity, combustion, and
    electromagnetism.
\end{abstract}

@@

\section{Introduction}
Implementing and maintaining robust and scalable applications is highly
relevant in the field of computational science and engineering. The efficient
production of such applications requires a rich ecosystem of state-of-the-art
reusable libraries \cite{Bartlett:2017:XFT}. They can then be used by domain
specialists who can decide what are the best tools available for their
applications. In this paper, we focus on the interoperability of two such
libraries.

@@

% Htool ?
Here, we describe the integration of the HPDDM solvers suite consisting of advanced iterative
methods and robust domain decomposition preconditioners into PETSc.
The design principles of the interface are showcased considering % alongside
minimalist and easy-to-run examples, with a focus on fulfilling reproducibility requirements as advocated by the
Association for Computing Machinery
\url{https://www.acm.org/publications/policies/artifact-review-badging}, as well as to 
ease the integration of these algorithms into more advanced frameworks
and facilitate the analysis and comparison of performance results~\cite{hoefler2015scientific}.
Furthermore, the benefits of code refactorization are shown for
applications that deal with multiple right-hand sides. This is the case of some
eigensolvers in SLEPc~\cite{hernandez2005ssf}, the Scalable Library for
Eigenvalue Problem computations. SLEPc is an add-on library that extends PETSc
with several classes for linear and nonlinear eigenvalue problems, as well as with
tools to facilitate the use of shift-and-invert spectral transformations.

@@

@@

All results have been obtained using double-precision arithmetic and 32-bit integers. The scaled efficiency measured in GFLOP per second is reported

@@

\subsubsection{Blocking inside SLEPc\label{sec:lobpcg}}
The implementations of the Locally Optimal Block Preconditioned Conjugate
Gradient~\cite{knyazev2001toward}~(LOBPCG) and the Contour Integral Spectrum
Slicing method~\cite{sakurai2003projection}~(CISS) in SLEPc were previously not
using blocking when applying the preconditioner or solving linear systems. In
version 3.14.0, these solvers have been adapted to employ \pk{KSPMatSolve} and
\pk{PCMatApply}.  In the case of LOBPCG, which is a purely blocked method, all
steps were already implemented as block operations, except the applications of
the preconditioner, which undermined the benefits of blocking. With the new
implementation, the gains are striking, as reported in~\cref{sec:lobpcg}.

@@

In this paper, we presente the interface between PETSc and HPDDM. The new
registered Krylov and overlapping Schwarz methods are explained and we show the
applicability of these  with numerical examples. Solving well-known problems,
we describe various runtime options. %Some more unconventional test
%cases were commented as well, for example using \pk{MatKAIJ} or \pk{MatIS} matrix types.
Overall, the interface paves the way for having recycling and block Krylov
methods plus robust overlapping Schwarz methods using the GenEO framework in
PETSc.

@@

Concerning SLEPc, we plan to implement new solvers or extend existing ones to
further exploit block solve primitives, for instance, block
Krylov--Schur~\cite{zhou2008block}. The current code for Arnoldi iterations
would then handle block Arnoldi iterations.

Concerning HPDDM, its integration inside PETSc offers a much-needed flexibility
compared to the standard implementation. Indeed, it is now possible to switch at
runtime many parameters that are determined at compile-time in the standard implementation, e.g.,
subdomain solvers and matrix types.

The interface between PETSc and HPDDM has still room for future improvements,
and we discuss openly two of them next. First, HPDDM handles mixed-precision,
e.g., using single-precision scalars for assembling coarse operators while
using double-precision scalars at the fine level. The topic of mixed-precision
is under scrutiny for the next major overhaul of PETSc. Second, HPDDM does not
currently handle GPU efficiently, but the interface with PETSc provides a very
thorough test bed with \pk{Mat} and \pk{PC} implementations that are exploiting
GPU.

@@

\bibitem{ambartsumyan2020hierarchical}
{\sc I.~Ambartsumyan, W.~Boukaram, T.~Bui-Thanh, O.~Ghattas, D.~Keyes,
  G.~Stadler, G.~Turkiyyah, and S.~Zampini}, {\em {Hierarchical Matrix
  Approximations of Hessians Arising in Inverse Problems Governed by PDEs}},
  arXiv:2003.10173,  (2020).

@@

\bibitem{anderson2019mfem}
{\sc R.~Anderson, J.~Andrej, A.~Barker, J.~Bramwell, J.-S. Camier, J.~Cerveny,
  V.~Dobrev, Y.~Dudouit, A.~Fisher, T.~Kolev, W.~Pazner, M.~Stowell, V.~Tomov,
  I.~Akkerman, J.~Dahm, D.~Medina, and S.~Zampini}, {\em {MFEM: A modular
  finite element methods library}}, Computers \& Mathematics with Applications,
   (2020), \url{http://mfem.org}.

@@

\bibitem{zhou2008block}
{\sc Y.~Zhou and Y.~Saad}, {\em {Block Krylov--Schur method for large symmetric
  eigenvalue problems}}, Numerical Algorithms, 47 (2008), pp.~341--359.
